Distance Sampling in \proglang{R}

### 2016-06-08 paper.Rnw
merge appendix A

>>>>

\appendix

\section*{Appendix A: Feature comparison}

There are four packages available for analysis of distance sampling data in \proglang{R} that we are aware of. All are available on CRAN. On top of \pkg{Distance} and \pkg{mrds} \citep{mrds-pkg} described above, they are \pkg{Rdistance} \citep{Rdistance-pkg} and \pkg{unmarked} \citep{Fiske:2011gv}. Table \ref{app-comp-tab} provides a feature comparison of these packages.

\begin{table}
\begin{tabular}{ccccc}
Feature    &  \pkg{Distance} & \pkg{Rdistance} & \pkg{unmarked} & \pkg{mrds}\\
\hline
Line transects & x & x & x & x\\
Point transets & x & & x & x\\
Interval (binned) distances & x & & x & x\\
Exact distances & x & x & & x\\
Continuous individual level covariates & x &  & & x\\
Factor individual level covariates & x & x & & x\\
Transect level covariates & x &  & x & x\\
Objects in clusters & x & x & x & x\\
Left truncation & x  & x & & x\\
Half-normal key & x & x & x & x\\
Hazard-rate key & x & x &  x & x\\
Uniform key & x & x & x & x\\
Gamma key &  & x &  & x\\
Negative exponential key &  &  & x  &\\
Adjustment terms & x & x & & x\\
AIC adjustment selection & x & x & \\
Monotonicity constraints & x &  & & x\\
Availability bias model & & & x &\\
Perception bias model & & & & x\\
Abundance estimation & x & x & x & x\\
Density estimation & x & x & x & x\\
User-defined likelihood functions & & x & & \\
\end{tabular}
\caption{Feature comparison of the available packages to perform distance sampling analyses.}
\label{app-comp-tab}
\end{table}

\subsection*{Abundance estimation}

\pkg{Distance} and \pkg{mrds} models can be used as part of a density surface model using \pkg{dsm}, which allows abundance to be modelled as a function of spatially varying covariates (such as location, sea depth, altitude etc). See \cite{Miller:2013fq} for more information.

\pkg{unmarked} also allows abundance to vary according to covariates, via the abundance part of the likelihood. See \cite{Fiske:2011gv} for more information on the package and \cite{Royle:2004cd} for more information on methodology.

Results \pkg{Rdistance} models can be use in combination with \proglang{R} modelling functions such as \code{lm}, \code{glm} etc to build abundance estimates which vary according to covariates. More information is available on the project's wiki.
